Exploratory data structure comparisons: Three new visual tools based on Principal Component Analysis

### 2018-06-17 article/Old versions/MPS/PCADSC_article.tex
New versions of all figures after update of anglePlot
Started editing of manuscript
CE-plot section updated

\usepackage[english]{babel}
\usepackage{authblk}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
%\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\usepackage{apalike}
\usepackage[utf8x]{inputenc}
%\usepackage{nameref,hyperref}
% line numbers
%\usepackage[right]{lineno}
% ligatures disabled
\usepackage{microtype}
\usepackage{multirow}
\usepackage[table]{xcolor}
\definecolor{lightgray}{HTML}{e6e6e6}

@@

The tools presented here are all based on comparing the PCA results across two different datasets that contain the same variables, $\mathbf{X}$ and $\mathbf{Y}$. Let 
$$\mathbf{Z} = \begin{pmatrix} \mathbf{X} \\ \mathbf{Y} \end{pmatrix} \in \RR^{n_x + n_y \times d}$$
 be the combined dataset. For each of these three datasets, we complete the following steps (here described for $\mathbf{X}$ only):
\begin{enumerate}
\item Standardize each of the variables to have mean zero and unit standard deviation. Let $\tilde{\mathbf{X}}_{j} \in \RR^{n_x \times d}$ be the standardized dataset.
\item Form the principal component analyses
\begin{align*}
S_x &= \frac{1}{n_x-1} \sum_{i=1}^{n_x} \tilde{x}_{i \cdot} \tilde{x}_{i \cdot}^\top = \sum_{j=1}^d \lambda_{j}^x \eta_{j}^x ({\eta_{j}^x})^\top 
\end{align*}
thereby obtaining loadings $\eta_{j}^x$ and eigenvalues $\lambda^x_j$ for $j=1,\dotsc,d$.
\end{enumerate}
The hereby obtained PCA decompositions of the correlation matrices can then be compared. The standardization implies that the diagonal elements of $S_x$, $S_y$, and $S_z$ all equal 1, and thus also that
\begin{equation*}
\sum_{j=1}^d \lambda_j^x = \sum_{j=1}^d \lambda_{j}^y = \sum_{j=1}^d \lambda_{j}^z =  d.
\end{equation*}
This identity will simplify some expressions below. Using a greedy interpretation of PCA, which is often overlooked in the literature, we can note that the sequence of loadings $\eta_j$ and their associated eigenvalues $\lambda_j$ yield a simultaneous description of the structure of the dataset for all approximating dimensions $q$. This implies that the loadings and the eigenvalues can be used to investigate the structure of the dataset without the need to decide on an approximating dimension, $q$, a priori.

@@

In \textit{Model 1} the observed variables $V_3$ and $V_4$ are caused by a common, unobserved variable $L_1$. Further details on the models and the simulated data are available in the supplementary material, including covariance matrices for the two models, in Appendix S2. In dataset $\mathbf{A}$, all observations stem from the same, underlying data structure and thus, the PCADSC plots should point towards no notable differences in data structures. In dataset $\mathbf{B}$, on the other hand, there are two fundamentally different, underlying data structures and therefore, the PCADSC plots should illustrate this lack of homogeneity in data structures.

@@

\subsection*{The cumulative eigenvalue plot}
The cumulative eigenvalue (CE) plot compares the eigenvalues of the correlation matrices. These eigenvalues represent how much information is withheld in each component in terms of explained variance. Thus, by comparing cumulative sums of eigenvalues, it is possible to obtain a detailed picture of how the two datasets differ in terms of what components are the most informative. In order to investigate whether the same proportion of the total variation can be described by the same number of principal components in the two datasets, we plot a piecewise linear curve connecting the points
\begin{align*}
&(0,0), \;
(\lambda_1^z,\lambda_{1}^x-\lambda_{1}^y), \;
(\lambda_1^z + \lambda_2^z,\lambda_{1}^x+\lambda_{2}^x-\lambda_{1}^y-\lambda_{2}^y), \;
\ldots, \\
&\bigg( \sum_{j=1}^d \lambda_j^z, \sum_{j=1}^d \lambda_{j}^x - \sum_{j=1}^d \lambda_{j}^y \bigg)
\end{align*}
Due to the standardization, the last point will always be equal to $(d,0)$. Thus, this curve will begin and end at the x-axis. And the larger excursions it makes away from the  first axis, the less alike the cumulative eigenvalue sums for the two datasets are. Moreover, a positive cumulative difference implies that dataset $\mathbf{X}$ holds more information in the first components than dataset $\mathbf{Y}$ does. By using the cumulative eigenvalues of $\mathbf{Z}$ as the first coordinate, we obtain a visualization where a component take up an amount of horizontal space that corresponds to the variance explained by the component (that is, its eigenvalue), thereby letting the more infuential components drive the visual impression.

@@

Two examples of cumulative eigenvalue plots are available in Figure \ref{plot.simCE}. In the top panel, we see a cumulative eigenvalue curve that falls well within the confidence region, thereby suggesting no difference in eigenvalues (as expected). This also agrees with the large $p$-values and the fact that only a single underlying model was used for generating this data. In the bottom panel, the conclusion is reversed, which is also in accordance with the nature of the simulated data, which contains two distinct groups. In order to test whether these cumulative differences are statistical artefacts or if they represent something real, we have implemented both the \emph{Kolmogorov-Smirnov} and the \emph{Cram\'er-von Mises} test statistics, which are given by
\begin{align*}
\text{KS} &= \max_{k=1,\dotsc,d} \bigg\lvert \sum_{j=1}^k \lambda_{j}^x - \sum_{j=1}^k \lambda_{j}^y \bigg\rvert, &
\text{CvM} &= \sum_{k=1}^{d-1} \frac{\lambda_k^z + \lambda_{k+1}^z}{2} {\bigg( \sum_{j=1}^k \lambda_{j}^x - \sum_{j=1}^k \lambda_{j}^y \bigg)}^2.
\end{align*}
We conduct the tests as \textit{permutation tests}, that is, by randomly reallocating the combined and separately standardized datasets into two new datasets of $n_x$ and $n_y$ observations, respectively, and then redoing the CE plot steps and recalculating the test statistics. This should be done a large (e.g. 10000) number of times. Then, a $p$-value is obtained by computing the proportion of reallocated datasets that lead to %even larger
test statistics at least as large as %than
the one we found for the original datasets.

@@

\subsection*{The angle plot}
The angle plot compares both loadings and eigenvalues at once and it can be used to understand the information loss if the data structure of one dataset is superimposed on the other, thereby revealing which principal components (i.e. loading and eigenvalue pairs) are most similar and most different across the two datasets. Let $\lambda_{\max} = \max\{ \lambda_{1}^x, \lambda_{1}^y \}$ be the largest eigenvalue for the two datasets. Then the empirical correlation matrix for the $\mathbf{X}$ dataset, $S_x$, has the following orthogonal decomposition in the coordinate system of the $\mathbf{Y}$ dataset
\begin{equation*}
S_x = \sum_{j=1}^d \lambda_{j}^x \eta_{j}^x ({\eta_{j}^x})^\top
= \lambda_{\max} \sum_{k=1}^d
\Bigg( \sum_{j=1}^d \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \eta_{j}^y ({\eta_{j}^y}^\top \eta_{k}^x) \Bigg)
{\Bigg( \sum_{j=1}^d \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \eta_{j}^y ({\eta_{j}^y}^\top \eta_{k}^x) \Bigg)}^\top,
\end{equation*}
and we have a similar decomposition of $S_y$ in the coordinate system of the $\mathbf{X}$ dataset. We propose to visualize these two decompositions in a $d \times d$ grid display. In the $j$th row and $k$th column of this display we plot two arrows based at the lower left corner of the grid cell. The first arrow has length $\mu_{jk}$ and angle $\theta_{jk}/2$ counterclockwise from the diagonal, and the second arrow has length $\nu_{jk}$ and angle $\theta_{jk}/2$ clockwise from the diagonal. To facilitate the following description we will refer to the arrows drawn counterclockwise as the blue arrows, and the arrows drawn clockwise as the red arrows. The lengths $\mu_{jk}$ and $\nu_{jk}$, and the angle $\theta_{jk}$, are given by
\begin{align*}
\mu_{jk} &= \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \lvert {\eta_{k}^x}^\top \eta_{j}^y \rvert, &
\nu_{jk} &= \sqrt{\frac{\lambda_{j}^y}{\lambda_{\max}}} \lvert {\eta_{j}^y}^\top \eta_{k}^x \rvert, &
\theta_{jk} &= \acos(\lvert {\eta_{k}^x}^\top \eta_{j}^y \rvert).
\end{align*}
For two $d$-dimensional, unit length vectors $a$ and $b$, it holds that $ a^\top b = \langle a, b \rangle = \tilde{c}(a,b)$, where $\tilde{c}$ denotes the empirical correlation function. Thus, in the angle plot, we are essentially looking at the absolute values of correlations between loadings that have been scaled according to their variance contributions. The absolute value of the projection $\eta_{k}^x {\eta_{j}^y}^\top$ is inserted due to the indeterminacy of the direction of loading vectors. This indeterminacy implies that the angle between loadings from the two datasets always can be chosen to be in the interval $[0,\pi/2]$, and hence the decomposition of $S_x$ and $S_y$ can be visualized in a joint plot by dividing the angles by two and using counterclockwise and clockwise shifts from the diagonal. Furthermore, the scaling of the lengths by $\lambda_{\max}$ is made so that the longest arrow has at most unit length. Figure \ref{plot.simAngle} presents two examples of angle plots based on simulated data.

@@

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure2_v2.pdf}
\caption{CE plots for the simulated datasets. Dataset A (with no differences in data structure) in top panel, and dataset B (where observations stem from different models) in bottom panel. Plots are annotated with the $p$-values of the Kolmogorov-Smirnov (KS) and the Cram\'er-von Mises (CvM) tests of the null hypothesis of no difference in data structures. The bold, black lines illustrate the observed cumulative eigenvalue differences, while the shaded areas are 95 \% confidence bands under the null hypothesis of no difference in data structures.}
\label{plot.simCE} 
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure3_v2.pdf}
\caption{Angle plots for the simulated datasets. Dataset A in top panel and dataset B in bottom panel. Blue arrows show the principal components of the observations in group 1 decomposed in the coordinate system of the principal components of group 2, while the red arrows illustrate the reverse.}
\label{plot.simAngle}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure4_v2.pdf}
\caption{Chroma plots for the simulated datasets. Dataset $\mathbf{A}$ in top panel and dataset $\mathbf{B}$ in bottom panel. Each bar represents a principal component and the bars are annotated with their cumulative percentages of explained variance.}
\label{plot.simChroma}
\end{figure}

@@

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure6_v2.pdf}
\caption{CE plots (top panel) and angle plots (bottom panel) comparing Danish and Bulgarian data on psychological well-being. The CE plot is annotated with the $p$-values of the Kolmogorov-Smirnov and the Cram\'er-von Mises tests of the hypothesis of no difference in data structures. In the angle plot, the blue arrows show the principal components of the Bulgarian dataset decomposed in the coordinate system of the principal components of the Danish dataset, while the red arrows illustrate the reverse.}
\label{plotBG.cehair}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure7_v2.pdf}
\caption{Chroma plot for comparing Danish and Bulgarian data on psychological well-being. A chroma plot comparing the 2nd, 3rd, and 4th principal components of the Bulgarian- and Danish psychological well-being data. The component-bars are annotated with their relative variance contributions. }
\label{plotBG.pancake}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure8_v2.pdf}
\caption{A CE plot (top panel) and an angle plot (bottom panel) comparing Danish and Swedish data on psychological well-being. The blue arrows show the principal components of the Danish dataset decomposed in the coordinate system of the principal components of the Swedish dataset, and the red arrows illustrate the reverse.}
\label{plotSE.cehair}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure9_v2.pdf}
\caption{Chroma plot comparing the loading patterns of the Danish and the Swedish subsamples. For each component, the bar is annotated with its cumulative variance contribution, that is, how much variance can be explained by having information of this and the preceding components.}
\label{plotSE.pancake}
\end{figure}

>>>>

\usepackage[english]{babel}
\usepackage{authblk}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
%\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\usepackage{apalike}
%\usepackage[utf8x]{inputenc}
\usepackage[utf8]{inputenc}
%\usepackage{nameref,hyperref}
% line numbers
%\usepackage[right]{lineno}
% ligatures disabled
\usepackage{microtype}
\usepackage{multirow}
\usepackage[table]{xcolor}
\definecolor{lightgray}{HTML}{e6e6e6}

@@

The tools presented here are all based on comparing the PCA results across two different datasets that contain the same variables, $\mathbf{X}$ and $\mathbf{Y}$. Let 
$$\mathbf{Z} = \begin{pmatrix} \mathbf{X} \\ \mathbf{Y} \end{pmatrix} \in \RR^{(n_x + n_y) \times d}$$
 be the combined dataset. For each of these three datasets, we complete the following steps (here described for $\mathbf{X}$ only):
\begin{enumerate}
\item Standardize each of the variables to have mean zero and unit standard deviation. Let $\tilde{\mathbf{X}}_{j} \in \RR^{n_x \times d}$ be the standardized dataset.
\item Form the principal component analyses
\begin{align*}
S_x &= \frac{1}{n_x-1} \sum_{i=1}^{n_x} \tilde{x}_{i \cdot} \tilde{x}_{i \cdot}^\top = \sum_{j=1}^d \lambda_{j}^x \eta_{j}^x ({\eta_{j}^x})^\top 
\end{align*}
thereby obtaining loadings $\eta_{j}^x$ and eigenvalues $\lambda^x_j$ for $j=1,\dotsc,d$.
\end{enumerate}
The hereby obtained PCA decompositions of the correlation matrices can then be compared. The standardization implies that the diagonal elements of $S_x$, $S_y$, and $S_z$ all equal 1, and thus also that
\begin{equation*}
\sum_{j=1}^d \lambda_j^x = \sum_{j=1}^d \lambda_{j}^y = \sum_{j=1}^d \lambda_{j}^z =  d.
\end{equation*}
This identity will simplify some expressions below. We note that the sequence of loadings $\eta_j$ and their associated eigenvalues $\lambda_j$ yield a simultaneous description of the structure of the dataset for all approximating dimensions $q$. This implies that the loadings and the eigenvalues can be used to investigate the structure of the dataset without the need to decide on an approximating dimension, $q$, a priori.

@@

In \textit{model 1} the observed variables $V_3$ and $V_4$ are caused by a common, unobserved variable $L_1$. Further details on the models and the simulated data are available in the supplementary material, including covariance matrices for the two models, in Appendix S2. In dataset $\mathbf{A}$, all observations stem from the same, underlying data structure and thus, the PCADSC plots should point towards no notable differences in data structures. In dataset $\mathbf{B}$, on the other hand, there are two fundamentally different, underlying data structures and therefore, the PCADSC plots should illustrate this lack of homogeneity in data structures.

@@

\subsection*{The cumulative eigenvalue plot}
The cumulative eigenvalue (CE) plot compares the eigenvalues of the correlation matrices. These eigenvalues represent how much information is withheld in each component in terms of explained variance. Thus, by comparing cumulative sums of eigenvalues, it is possible to obtain a detailed picture of how the two datasets differ in terms of what components are the most informative. In order to investigate whether the same proportion of the total variation can be described by the same number of principal components in the two datasets, we plot a piecewise linear curve connecting the points
\begin{align*}
&(0,0), \;
(\lambda_1^z,\lambda_{1}^x-\lambda_{1}^y), \;
(\lambda_1^z + \lambda_2^z,\lambda_{1}^x+\lambda_{2}^x-\lambda_{1}^y-\lambda_{2}^y), \;
\ldots, \\
&\bigg( \sum_{j=1}^d \lambda_j^z, \sum_{j=1}^d \lambda_{j}^x - \sum_{j=1}^d \lambda_{j}^y \bigg)
\end{align*}
Due to the standardization, the last point will always be equal to $(d,0)$. Thus, this curve will begin and end at the x-axis. And the larger excursions it makes away from the  first axis, the less alike the cumulative eigenvalue sums for the two datasets are. Moreover, a positive cumulative difference implies that dataset $\mathbf{X}$ holds more information in the first components than dataset $\mathbf{Y}$ does. By using the cumulative eigenvalues of $\mathbf{Z}$ as the first coordinate, we obtain a visualization where a component take up an amount of horizontal space that corresponds to the variance explained by the component (that is, its eigenvalue), thereby letting the more influential components drive the visual impression.

@@

In order to test whether these cumulative differences are statistical artefacts or if they represent something real, we have implemented both the \emph{Kolmogorov-Smirnov} and the \emph{Cram\'er-von Mises} test statistics, which are given by
\begin{align*}
\text{KS} &= \max_{k=1,\dotsc,d} \bigg\lvert \sum_{j=1}^k \lambda_{j}^x - \sum_{j=1}^k \lambda_{j}^y \bigg\rvert, &
\text{CvM} &= \sum_{k=1}^{d-1} \frac{\lambda_k^z + \lambda_{k+1}^z}{2} {\bigg( \sum_{j=1}^k \lambda_{j}^x - \sum_{j=1}^k \lambda_{j}^y \bigg)}^2.
\end{align*}
We conduct the tests as \textit{permutation tests}, that is, by randomly reallocating the combined and separately standardized datasets into two new datasets of $n_x$ and $n_y$ observations, respectively, and then redoing the CE plot steps and recalculating the test statistics. This should be done a large (e.g. 10000) number of times. Then, a $p$-value is obtained by computing the proportion of reallocated datasets that lead to %even larger
test statistics at least as large as %than
the one we found for the original datasets.

@@

Two examples of cumulative eigenvalue plots are available in Figure \ref{plot.simCE}. In the top panel, we see a cumulative eigenvalue curve that falls well within the confidence region, thereby suggesting no difference in eigenvalues (as expected). This also agrees with the large $p$-values and the fact that only a single underlying model was used for generating this data. In the bottom panel, the conclusion is reversed, which is also in accordance with the nature of the simulated data, which contains two distinct groups.

\subsection*{The angle plot}
The angle plot compares both loadings and eigenvalues at once and it can be used to understand the information loss if the data structure of one dataset is superimposed on the other, thereby revealing which principal components (i.e. loading and eigenvalue pairs) are most similar and most different across the two datasets. Let $\lambda_{\max} = \max\{ \lambda_{1}^x, \lambda_{1}^y \}$ be the largest eigenvalue for the two datasets. Then the empirical correlation matrix for the $\mathbf{X}$ dataset, $S_x$, has the following orthogonal decomposition in the coordinate system of the $\mathbf{Y}$ dataset
\begin{equation*}
S_x = \sum_{j=1}^d \lambda_{j}^x \eta_{j}^x ({\eta_{j}^x})^\top
= \lambda_{\max} \sum_{k=1}^d
\Bigg( \sum_{j=1}^d \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \eta_{j}^y ({\eta_{j}^y}^\top \eta_{k}^x) \Bigg)
{\Bigg( \sum_{j=1}^d \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \eta_{j}^y ({\eta_{j}^y}^\top \eta_{k}^x) \Bigg)}^\top,
\end{equation*}
and we have a similar decomposition of $S_y$ in the coordinate system of the $\mathbf{X}$ dataset. We propose to visualize these two decompositions in a $d \times d$ grid display. In the $j$th row and $k$th column of this display we plot two arrows based at the lower left corner of the grid cell. The first arrow has length $\mu_{jk}$ and angle $\theta_{jk}/2$ counterclockwise from the diagonal, and the second arrow has length $\nu_{jk}$ and angle $\theta_{jk}/2$ clockwise from the diagonal. To facilitate the following description we will refer to the arrows drawn counterclockwise as the blue arrows, and the arrows drawn clockwise as the red arrows. The lengths $\mu_{jk}$ and $\nu_{jk}$, and the angle $\theta_{jk}$, are given by
\begin{align*}
\mu_{jk} &= \sqrt{\frac{\lambda_{k}^x}{\lambda_{\max}}} \lvert {\eta_{k}^x}^\top \eta_{j}^y \rvert, &
\nu_{jk} &= \sqrt{\frac{\lambda_{j}^y}{\lambda_{\max}}} \lvert {\eta_{j}^y}^\top \eta_{k}^x \rvert, &
\theta_{jk} &= \acos(\lvert {\eta_{k}^x}^\top \eta_{j}^y \rvert).
\end{align*}
For two $d$-dimensional, unit length vectors $a$ and $b$, it holds that $ a^\top b = \langle a, b \rangle = \tilde{c}(a,b)$, where $\tilde{c}$ denotes the empirical correlation function. Thus, in the angle plot, we are essentially looking at the absolute values of correlations between loadings that have been scaled according to their variance contributions. The absolute value of the projection ${\eta_{k}^x}^\top \eta_{j}^y$ is inserted due to the indeterminacy of the direction of loading vectors. This indeterminacy implies that the angle between loadings from the two datasets always can be chosen to be in the interval $[0,\pi/2]$, and hence the decomposition of $S_x$ and $S_y$ can be visualized in a joint plot by dividing the angles by two and using counterclockwise and clockwise shifts from the diagonal. Furthermore, the scaling of the lengths by $\lambda_{\max}$ is made so that the longest arrow has at most unit length. Figure \ref{plot.simAngle} presents two examples of angle plots based on simulated data.

@@

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure2_v3.pdf}
\caption{CE plots for the simulated datasets. Dataset A (with no differences in data structure) in top panel, and dataset B (where observations stem from different models) in bottom panel. Plots are annotated with the $p$-values of the Kolmogorov-Smirnov (KS) and the Cram\'er-von Mises (CvM) tests of the null hypothesis of no difference in data structures. The bold, black lines illustrate the observed cumulative eigenvalue differences, while the shaded areas are 95 \% confidence bands under the null hypothesis of no difference in data structures.}
\label{plot.simCE} 
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure3_v3.pdf}
\caption{Angle plots for the simulated datasets. Dataset A in top panel and dataset B in bottom panel. Blue arrows show the principal components of the observations in group 1 decomposed in the coordinate system of the principal components of group 2, while the red arrows illustrate the reverse.}
\label{plot.simAngle}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure4_v3.pdf}
\caption{Chroma plots for the simulated datasets. Dataset $\mathbf{A}$ in top panel and dataset $\mathbf{B}$ in bottom panel. Each bar represents a principal component and the bars are annotated with their cumulative percentages of explained variance.}
\label{plot.simChroma}
\end{figure}

@@

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure6_v3.pdf}
\caption{CE plots (top panel) and angle plots (bottom panel) comparing Danish and Bulgarian data on psychological well-being. The CE plot is annotated with the $p$-values of the Kolmogorov-Smirnov and the Cram\'er-von Mises tests of the hypothesis of no difference in data structures. In the angle plot, the blue arrows show the principal components of the Bulgarian dataset decomposed in the coordinate system of the principal components of the Danish dataset, while the red arrows illustrate the reverse.}
\label{plotBG.cehair}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure7_v3.pdf}
\caption{Chroma plot for comparing Danish and Bulgarian data on psychological well-being. A chroma plot comparing the 2nd, 3rd, and 4th principal components of the Bulgarian- and Danish psychological well-being data. The component-bars are annotated with their relative variance contributions. }
\label{plotBG.pancake}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure8_v3.pdf}
\caption{A CE plot (top panel) and an angle plot (bottom panel) comparing Danish and Swedish data on psychological well-being. The blue arrows show the principal components of the Danish dataset decomposed in the coordinate system of the principal components of the Swedish dataset, and the red arrows illustrate the reverse.}
\label{plotSE.cehair}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[scale=0.8]{Figure9_v3.pdf}
\caption{Chroma plot comparing the loading patterns of the Danish and the Swedish subsamples. For each component, the bar is annotated with its cumulative variance contribution, that is, how much variance can be explained by having information of this and the preceding components.}
\label{plotSE.pancake}
\end{figure}
